Social inequality in mortality among adults and elderly in northern Sweden 1851--2013\footnote{Paper presented at the session 'Health and Mortality in Arctic Populations'. The Ninth International Congress on Arctic Social Sciences, Umeå, June 8--12, 2017}

### 2017-05-30 icass.Rnw
Revising discussion. Adding references

\title{Social inequality in mortality among
  adults and elderly in northern Sweden 1851--2013\footnote{Paper presented
    at the
    session 'Health and Mortality in Arctic Populations'. The Ninth International
    Congress on Arctic Social Sciences,
    Umeå,
    June 8--12, 2017}}

\date{May 29, 2017}

@@

The classification scheme we use is  based on HISCLASS \citep{lema11}, a class-based system. The scheme uses HISCO for assigning categories. HISCO (Historical International Standard Classifications of Occupations) \citet{lema02}  is a classification system of historical occupations, categorising occupations according to tasks and skills. We have merged the original 12 classes in HISCLASS, based on position on the labour market, ending up with the following three and an additional fourth category with unknown class:

\begin{enumerate}
\item {\tt elite}, HISCLASS 1 and 2. Higher managers and professionals.
\item {\tt middle}, HISCLASS 3, 4, 5, 6 and 8. Lower managers and farmers.
\item {\tt worker}, HISCLASS 7, 9, 10, 11 and 12. Workers of different skills including farm workers.
%%\item Lower skilled farm workers, HISCLASS 10 and 12, Lower and unskilled farm workers
\end{enumerate}

@@

@@

@@

A common alternative to a class analysis is to use income or education for investigating socio-economic health differences. 
Comparing different approaches are furthermore valuable in order to understand why some are advantaged and others 
disadvantaged. For the time being, we have to restrict such a comparative analysis to the last decades.

@@

The WHO Commission on Social Determinants of Health framework for the social determinants of health discusses how socio-economic position impacts on health. The influences are not direct but works through intermediate determinants related to exposure and vulnerability \citep{who10}. People are placed within socio-economic and political contexts, as well as value systems in the society. The socio-economic position imply different access to resources, power and prestige, all aspects of crucial importance. The framework takes a holistic approach to the social determinants, including not only the effect of pure economic and material resources. The importance of understanding people in their context has also been emphasised by for example \citet{kri05}. What she calls embodiment refers to the recognition that human beings are simultaneously social beings and biological organisms.

The authors of this paper have scrutinised different aspects of the social
pattern of mortality in northern Sweden for quite some time, focusing on
different regions, different age groups and partly different social
categories. Already the earliest findings, concerning the social inequality
in a rapidly growing town in an industrial district, surprised and at least
at first confused the author \citep{edv92}. The expected social gradient in
survival, meaning that people from higher social classes always had an
advantage, was in many cases, particularly among adult men, not found
despite the well-documented socio-economic differences in this
society. There were however large differences between men and women and for
different age groups. Further studies have confirmed these
findings \citep{edli11, edbr12}. These earlier studies have ended with an
assumption that distinct and consistent differences to the advantage of higher
classes developed somewhere during and probably quite early in the 20th
century. Until now we have however not been able to investigate the
continuous development on mortality differences among the adult and elderly
population due to no available data. This is the first time it has been
possible, and the results continue to surprise us. The consistent advantage
for higher social classes do not appear until quite late in the regions we
have analysed. The social pattern of female mortality are closer to what is
expected, but a social gradient in male adult mortality seems to be a
recent phenomenon. Furthermore, this transition and the class pattern of mortality differ depending on age group. Inequality in mortality appear clearer for the population at working age, while the differences are smaller among the elderly.

We have also added analyses of cause-specific mortality to the previous version. Mortality from cardiovascular diseases contributes to a large proportion of all deaths during the 20th century, and we are not surprised that it has a fairly similar pattern as for all-cause mortality. What is surprising is however the small class differences among the elderly. Cancer is sometimes considered less connected to social class, and we find mixed results in our analyses. There are however indications of a social gradient during the last decades in both studied age groups. Development of death causes are however problematic since the ways they have been diagnosed and the proportion with missing or vague information has changed over time.

The perhaps surprising results makes it reasonable to consider their
validity and possible weaknesses in the study. There are of course
uncertainties involved in all studies of this sort, something that should
not be ignored.  One obstacle concerns the quality of the available
information in the sources. Occupational titles in historical sources are
not always distinct enough to define the correct class. It can be difficult
to identify the skill level in the working class population because many
were often only designated as labourer. The size of the enterprise for free
entrepreneurs is unknown and it may be impossible to distinguish if they
are self-employed or have many employed. The class system and the position
of different classes can also be discussed. Studies covering long period
have the problem that occupations and classes change character - they do
not represent the same thing. Nonetheless, these ambiguities and possible
mis-classifications are not of that sort and magnitude that it would radically change
the basic patterns observed. The social classes identified during the
different periods represent distinct groups with very large differences in
resources, capacities and power. A worker would not end up in the highest
SOCPO level and large-scale entrepreneurs and higher officials would
certainly be assigned the right level.Available data also show that the
differences in income and wealth were large, often enormous.

In this paper we have analysed social inequality in health from the
perspective of social classes using HISCLASS. However, there are
alternative ways to analyse social inequality, and we cannot exclude that
such analyses gives other results or at least would modify the image
presented here. There are other alternatives then HISCLASS, representing
status or class dimensions. SOCPO is another system based on hierarchic
categories of social power, which we have used in a couple of previous
studies. In many modern studies categorisations based on income, wealth
or education are common for studies of inequality. It is possible that
such categorisations are better fit to distinguish between groups in
society. They represent different dimensions of inequality, and we believe
that the topic discussed here would gain from analyses of different aspects
of inequality. One problem is however that reliable data on income and
education are rarely available in historical population databases. We can
however be certain that the highest HISCLASSES level have more education
and the lowest the least. We also know from the 1930 census that the
economic differences were very large at that time and that occupations in
higher HISCLASSES tend to have higher incomes. The groups that may diverge
are those within agriculture and those having a small enterprise where we
can assume that the reported income figures underestimate their economic
standard.

As a test, we have however included analyses of the associations of
education and income to mortality for the period 1990--2005. For women as
well as for men, for both age groups and for all-cause mortality as well as
cause-specific mortality we find consistent differences with higher
educational level and more income leading to lower mortality. As already
mentioned, the class analysis proved that it in most cases are strong
differences during the last centuries, but the pattern is much clearer in
the alternative ways of categorisation. This does not mean that a class
analysis is worthless, rather that the different categorisations measure
different things. But it certainly point to the importance to also consider
other aspect of social inequality in the past. We do not know if education
and income are equally important in the past and that these measures would
provide us with a completely different picture, but this is something that
calls for further investigation.

The results of small socio-economic gradient in the past and its late development are however in accordance with other studies \citep{bevp11},
for example by \citet{bedr11} in a very different setting, a couple of
Scanian parishes. They found the first indications of an emerging social
gradient in the 1950s.The similarities makes us more comfortable about our
results.

\citet{who10} lists three different types of intermediary social
determinants of health, material circumstances, behaviours and biological
factors, and finally psycho-social factors. Most studies have focused on
the material circumstances. Having a better economy can give advantages in
many aspects of relevance for health and mortality, for example good living
conditions, access to food, medical care, better conditions at work et
cetera. Even though these advantages ought to have been at hand in our
studied regions, the final result was not better health for men in the most
privileged groups until the later part of the 20th century. We must
consider other intermediary determinants. Psycho-social factors have been
frequently discussed during the most recent decades, for example by
\citet{mar04}. He argues that being higher in the status ladder leads to
better health, regardless of where in the status ladder we look. This is
also one of the arguments \citet{wipi09} use as an explanation why equal
societies always perform better than unequal ones when it comes to health
and other social conditions. In the case studied here, we do not find any
indication of a status syndrome, since mortality levels was in the opposite
directions to the one we would expect from these theories.

This makes us consider the possible impact of behavioural factors. Our results of the gender aspect might be of interest in this case. When it comes to gender differences, the present study confirms what we have found on the Sundsvall region. There is a clear gendered pattern when it comes to social inequality in mortality. While high HISCLASS did not have any positive impact on survival for men in the early 20th century, we find  that higher social
positions among women led to longer lives. Within the same households, the effect of social
class seems to lead to different results. This opens up for behavioural
aspects based on different gendered expectations. Alcohol has been
suggested as part of the large gender differences in the 19th century
\citep{wil99, edv92}, and a more risk-taking life may be a component of the
male mortality.

Having a high social status could be expressed in attitudes and behaviours
that were harmful for health. Alcohol consumption decreased rapidly during
the period late 19th century to the 1950s, thus possibly making room for an
increased role of economic resources. But the old pattern in social health
inequalities remained. It is however possible that the relative differences
in alcohol consumption prevailed, although at a lower level. The question
is however if the lower levels of alcohol consumption had any substantial
negative impact on health. Swedes at that time had very low consumption
levels. Alternative explanations should also be scrutinise. There
may have been strong and heavy requirements of being successful for the
male \"elite. Many were involved in businesses with high risks. It would be
very damaging to lose the high position and they may have felt that they
lived under constant threat of economic failure. There are several examples among
the economic \"elite in the Sundsvall region that they had such problems
\citep{edli11}. A possible future study involving analyses of death causes
might throw some light on this issue.

Our study illustrates the importance of considering contextual factors and
the restrictions of making general statements. We need to look into all
aspects of social classes, not only the material ones but also what this
means for behaviour, norms, way of living. \citet{edli11} suggested that an
explanation for the larger social health inequalities in the present-day
world is that status and higher social position is now expressed through
health and being fit and physical active, while in many historical contexts
higher social status was expressed by low physical activity, obsessive
eating (being fat was impressive), drinking, smoking and sometimes also
taking risks.

Another question is if our results can be generalised. We have studied
environments in northern Sweden without any large metropolis. Perhaps
Sweden and/or this part of Sweden is a special case. The social differences
may have been smaller, the involvement of government both at the national
and local level may have been stronger, thus mitigating socio-economic
disadvantages. These are relevant questions, which hopefully can be
addressed in other and maybe comparative studies.

Finally, we argue that our results have implications for our understanding
on several hot research issues. A central question concerning the mortality
transition is how the enormous increase in survival could be
accomplished. This is still a hot debated question. \citet{mck76} argues that the
improvements up to the middle of the 20th century was first of all made
possible by improved nutrition. This seems to implicate that having more
resources would lead to better survival. If research shows that the most
advantaged groups did not have better health and longer lives, this do not
necessarily speak against McKeown's ideas but it makes the question more
complex. The results are also difficult to fit into \citet{liph95}
hypothesis that social conditions should be seen as a fundamental cause of
disease. Even if we would eliminate disadvantages such as bad living
conditions, lower standard of medical care and other proximate determinants
for the poor, the health disadvantage would remain. In the historical
context we have analysed here, this does not get any support. This further
illustrate the need for situating the analyses in their historical and
geographical contexts.

In recent years, the question about the relation between health and income
inequality been widely discussed. Even though our study do not address this
issue directly, our results are of interest for this question as well. The
development of the social inequality in mortality do not seem to fit that
well into the development of income and wealth inequalities. The early 20th
century, in Sweden as well as in other countries, was characterised by very
large economic differences, still this did in our regions not lead to any
consistent advantage for the highest social classes. After the Second World
War, economic differences diminished, and was at its lowest in Sweden at
about 1980, i.e. at a time when a more expected pattern of a social
hierarchy in survival had appeared. The period thereafter is characterised
by increasing economic differences, and this may coincide with increasing
health differences as well. The long-term relation between economic
inequality and health is worth further studies.

\section{Conclusion}

We started this paper by raising the issue of the development of social
inequalities. When studying this question in history we often get surprised
of its complexity. In a society characterised by large economic inequality,
we would expect that this would result in corresponding mortality
differences, i.e. higher social classes having an advantage. We find the
appearance of a consistent social pattern in mortality, to the advantage of
the higher classes, but surprisingly late in our history. It is only in the
recent decades that this pattern dominates. Improved survival characterised
however all HISCLASSes. We can however confirm that there are gender
differences in the social pattern. While there is no clear social gradient
among men, the results at least partly indicate advantages for the higher
social classes for women already early in the 20th century. Females had
more to gain from a high social position. Previous studies of the mortality
transition have made similar observations. Among adults, gender has often
been of more significant importance than social class. The reason for this
difference and for the late appearance of the modern social pattern among
men is still an enigma that require further studies, even though we have
suggested the possible importance of behavioural factors and changes in
ideals connected to different social classes. This should stimulate us to
further disentangle the pathways for how social position influence health
and mortality in different societies. Furthermore, we do not find any substantial differences or distinct pattern in the relation between class and the two major cause-of-death groups. Finally, we find stronger differences in the social pattern among the working-age population, while the associations were smaller or not so clear among the elderly. However, when analysing the association with education and income in the period 1990--2005 we find consistent patterns of higher income and higher education level being associated with lower mortality, indicating that these variables are better to catch socio-economic differences at least in the present-day society.

>>>>

\title{Long--term perspectives of social inequality in mortality among
  adults and elderly in northern Sweden 1851--2013\footnote{Paper presented
    at the
    session 'Health and Mortality in Arctic Populations'. The Ninth International
    Congress on Arctic Social Sciences,
    Umeå,
    June 8--12, 2017}}

\date{May 30, 2017}

@@

HISCLASS, the classification system used here, is a ”. . . HISCO-based
historical international social class scheme” \citep{lema11, lema02}. The different classes in HISCLASS represent distinct categories, based on whether the work is manual or non–manual, skill level, supervision and sector. Implicitly it reflects large differences in what is in the forefront of other
categorisations, as for example access to economic and other resources, status, social power et cetera. We have merged the original 12 classes in HISCLASS, based on position in the labour market, ending up with the following three and an additional fourth category with unknown class:

\begin{enumerate}
\item {\tt elite}, HISCLASS 1 and 2. Higher managers, professionals and higher white collar.
\item {\tt middle}, HISCLASS 3, 4, 5, 6 and 8. Lower managers, farmers and lower white collar.
\item {\tt worker}, HISCLASS 7, 9, 10, 11 and 12. Workers including farm workers.
%%\item Lower skilled farm workers, HISCLASS 10 and 12, Lower and unskilled farm workers
\end{enumerate}

@@

\subsection{Cardiovascular mortality}

@@

\subsection{Cancer mortality}

@@

A common alternative to a class analysis is to use income or education for investigating socio-economic health differences. \citet{goldk10} criticize how many economists and epidemiologists operationalize inequality in their studies, for example by not distinguishing between class and status aspects. Class refers to relational aspects in society, while many other variables focus on the distribution of material and/or immaterial resources. He advocates studies of different aspects, which together complement each other. Comparing different approaches are valuable in order to understand why some are advantaged and others disadvantaged and what sort of factors determines health differences. In this and the following section, we introduce analyses of education and income. In that way, we can compare the class analyses with these alternative systems. For the time being, we have to restrict such a comparative analysis to the last decades.

@@

The authors of this paper have scrutinised different aspects of the social
pattern of mortality in historical northern Sweden for some time, focusing on
different regions, different age groups and different time periods \citep{edv92, edv04, edli11, edbr12}. When it comes to infant and child mortality pre-- and early--transitional time, the expected socio-economic differences was often but not always confirmed \citep{edv92, edv04}, something that has been found in other studies as well \citep{bredro02, bube01, mol17}. Studies based on microdata on the adult population in history have been rare, mainly due to lack of adequate data. The results from such studies show other patterns than that of infant and child mortality. The expected social gradient in
survival, meaning that people from higher social classes always had an
advantage, was in many cases, particularly among adult men, not found
despite the well-documented socio-economic differences \citep{edli11, edbr12}. There were however large differences between men and women and for different age groups. These earlier studies have suggested that distinct and consistent differences to the advantage of higher classes developed somewhere during and probably quite early in the 20th century \citep{bevp11}. Until now we have however not been able to investigate the continuous development on mortality differences among the adult and elderly
population due to no available data. This is the first time it has been possible, and the results continue to surprise us. The consistent advantage for higher social classes do not appear until quite late in the regions we have analysed. The social pattern of female mortality are in line to what is
expected, but a social gradient in male adult mortality seems to be a recent phenomenon. Furthermore, this transition and the class pattern of mortality differ depending on age group. Inequality in mortality appear clearer for the population at working age, while the differences are smaller among the elderly.

The results of small socio-economic gradient in the past and its late development are however in accordance with other studies \citep{bevp11}, for example by \citet{bedr11} in a very different setting, a couple of Scanian parishes, although with somewhat different social categories. They found the first indications of an emerging social gradient in the 1950s.

We have also added analyses of cause-specific mortality to the previous version. Mortality from cardiovascular diseases contributes to a large proportion of all deaths during the 20th century, and we are not surprised that it has a fairly similar pattern as for all-cause mortality. Cancer is sometimes considered less connected to social class, and we find mixed results in our analyses.

The surprising results makes it reasonable to consider the
validity and possible weaknesses in the study. One obstacle concerns the quality of available information. Occupational titles in historical sources are not always distinct enough to define the correct class. It can be difficult to identify the skill level in the working class population because many
were often only designated as labourer. The size of the enterprise for free
entrepreneurs is unknown and it may be impossible to distinguish if they
are self-employed or have employees. The class system and the position
of different classes can also be discussed. Studies covering long period
have the problem that occupations and classes change character - they do
not represent the same thing. Nevertheless, these ambiguities and possible
mis-classifications are not of that sort and magnitude that it would radically change
the basic patterns observed. The social classes identified during the
different periods represent distinct groups with very large differences in
resources, capacities and power. Available data, for example from the 1930 census confirm that class differences in income and wealth were large, often enormous.

In this paper we have analysed social inequality in health from the
perspective of social classes using HISCLASS. However, there are
alternative ways to analyse social inequality, and we cannot exclude that
such analyses gives other results or at least would modify the image
presented here. In many modern studies categorisations based on income, wealth
or education are common for studies of inequality. They represent different dimensions of 
inequality. One problem is however that reliable data on income and
education are rarely available in historical population databases. 
As a test, we have included analyses of the associations of
education and income to mortality for recent decades. For women as
well as for men, for both age groups and for all-cause mortality as well as
cause-specific mortality we find consistent differences with higher
educational level and more income leading to lower mortality, thus basically confirming what we find in the class analysis. We do not know if education and income are equally important in the past.

The WHO Commission on Social Determinants of Health framework for the social determinants of health provides a model on how socio-economic position influence health. Social class works through intermediate determinants related to exposure and vulnerability \citep{who10}. People are situated within socio-economic and political contexts, as well as prevailing value systems in society. Socio-economic position imply different access to resources, power and prestige, all being aspects of crucial importance. The framework takes a holistic approach to the social determinants, including not only the effect of pure economic and material resources. The importance of understanding people in their context is also emphasised by for example \citet{kri05}. What she calls embodiment refers to the recognition that human beings are simultaneously social beings and biological organisms. The WHO framework 
\citep{who10} lists three different types of intermediary social determinants of health; material circumstances, behaviours, biological factors, and finally psycho-social factors. Most studies focus on material circumstances. Having a good economy gives advantages in many aspects of relevance for health and mortality, for example good living conditions, access to food, medical care and better working conditions. Even though these advantages were at hand in our studied regions, the final result was not better health for men in the most privileged groups until the late 20th century. We must therefore consider other explanations. Psycho-social factors have been frequently discussed during the most recent decades, for example by \citet{mar04}, who argues that a higher status leads to better health, regardless of where in the status ladder we make our observations. Psychosocial factors are also suggested by \citet{wipi09} as an explanation why equal societies always perform better than unequal ones when it comes to health and other social conditions. In the case studied here, we do not find any indication of a status syndrome, since mortality levels was in the opposite directions to the one we would expect from these theories.

This makes us consider the possible impact of behavioural factors. Our results of the gender aspect might be of interest in this case. When it comes to gender differences, the present study confirms what we previously have found in the Sundsvall region \citep{edli11}. There is a clear gendered pattern when it comes to social inequality in mortality.  Within the same households, the effect of social class seems to lead to different results, high social position being beneficial for women but not for men. We might consider the role of behaviours based on different gendered expectations. Having a high social status could be expressed in attitudes and behaviours that were harmful for health (this is discussed in more detail in \citet{edli11}. Alcohol has been suggested as part of the large gender differences in the 19th century \citep{wil99, edv92}, and a more risk-taking life may be a component of male mortality \citep{cour00}. There are however some problems with attributing most of the male excess mortality to heavy drinking. Alcohol consumption decreased substantially during the period late 19th century to the 1950s, but the old pattern in social health inequalities remained. Swedes at that time had very low consumption levels. Alternative explanations should also be considered. There
were strong and heavy requirements of being successful for the male \"elite. Many were involved in businesses with high risks and they may have felt a constant threat of economic failure.

Our study illustrates the importance of considering contextual factors and
the restrictions of making general statements. We need to look into all
aspects of social classes, not only the material ones but also what this
means for behaviour, norms, way of living. \citet{edli11} suggest that an
explanation for the larger social health inequalities in present-day
society is that high status is expressed through health, being fit and physically active, while in many historical contexts higher social status (for men) was expressed in low physical activity, obsessive
eating, drinking, smoking and sometimes also taking risks. Different historical contexts promote different ways to spend economic resources.

One of the major findings in this study relate to gender. As several researchers have pointed out, demography has often ignored power relations and different roles within families \citep{watkins93, mack95, jans07}. By investigating possible differences in mortality patterns according to sex, we have identified gendered patterns within the family. Different patterns between the sexes may be caused by discrimination on the female part in the household. We know that the society was
strongly segregated according to gender and that women were discriminated
and had less space for action. In our case, we do not find this had negative consequences for women, indicating that the relation between sexessociety was characterised by distinct gendered expectations on behaviour.

Another question is if our results can be generalised to other historical contexts. We have studied
environments in northern Sweden without any large metropolis. Perhaps
Sweden and/or this part of Sweden is a special case. The social differences
may have been smaller, the involvement of government both at the national
and local level may have been stronger, thus mitigating socio-economic
disadvantages. In comparison with many other countries, the level of mortality has been fairly low. These are relevant questions, which hopefully can be
addressed in other and maybe comparative studies.

Finally, we argue that our results have implications for the understanding
on several other hot research topics. A central and hot debated question concerning the mortality
transition is how the enormous increase in survival was accomplished. \citet{mck76} argues that the
improvements up to the middle of the 20th century was primarily caused by improved nutrition. This implicate that having more resources would lead to better survival. If research shows that the most
advantaged groups did not have better health and longer lives, this do not
necessarily speak against McKeown's ideas but it makes the question more
complex. The results are also difficult to fit into \citet{liph95}
hypothesis that social conditions should be seen as a fundamental cause of
disease and not only expressions of different disadvantages. In the historical
context analysed here, this idea is not supported.

In recent years, the question about the relation between health and income
inequality been widely discussed \citep{wipi09}. Even though our study do not address this
issue directly, our results are of interest for this question as well. The
development of the social inequality in mortality do not seem to fit that
well into the development of income and wealth inequalities. The early 20th
century, in Sweden as well as in other countries, was characterised by very
large economic differences, still this did in our regions not lead to any
consistent advantage for the highest social classes. After the Second World
War, economic differences diminished, and was at its lowest in Sweden at
about 1980, i.e. at a time when a more expected pattern of a social
hierarchy in survival had appeared. The period thereafter is characterised
by increasing economic differences, and this may coincide with increasing
health differences as well. The long-term relation between economic
inequality and health is worth further studies.

% \section{Conclusion}
% 
% We started this paper by raising the issue of the development of social
% inequalities. When studying this question in history we often get surprised
% of its complexity. In a society characterised by large economic inequality,
% we would expect that this would result in corresponding mortality
% differences, i.e. higher social classes having an advantage. We find the
% appearance of a consistent social pattern in mortality, to the advantage of
% the higher classes, but surprisingly late in our history. It is only in the
% recent decades that this pattern dominates. Improved survival characterised
% however all HISCLASSes. We can however confirm that there are gender
% differences in the social pattern. While there is no clear social gradient
% among men, the results at least partly indicate advantages for the higher
% social classes for women already early in the 20th century. Females had
% more to gain from a high social position. Previous studies of the mortality
% transition have made similar observations. Among adults, gender has often
% been of more significant importance than social class. The reason for this
% difference and for the late appearance of the modern social pattern among
% men is still an enigma that require further studies, even though we have
% suggested the possible importance of behavioural factors and changes in
% ideals connected to different social classes. This should stimulate us to
% further disentangle the pathways for how social position influence health
% and mortality in different societies. Furthermore, we do not find any substantial differences or distinct pattern in the relation between class and the two major cause-of-death groups. Finally, we find stronger differences in the social pattern among the working-age population, while the associations were smaller or not so clear among the elderly. However, when analysing the association with education and income in the period 1990--2005 we find consistent patterns of higher income and higher education level being associated with lower mortality, indicating that these variables are better to catch socio-economic differences at least in the present-day society.
